Collaborative Spatial Reuse in Wireless Networks via Selfish Multi-Armed Bandits

### 2018-09-06 OLD SUBMISSION/Article (new version) - LaTeX files/collaborative_spatial_reuse_in_wireless_networks_via_selfish_mabs.tex
New + old submission differentiation

%% 
%% Copyright 2007, 2008, 2009 Elsevier Ltd
%% 
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%

%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%% SP 2008/03/01

\documentclass[preprint,12pt]{elsarticle}

\usepackage{graphicx} 	  %% The graphicx package provides the includegraphics command.
\usepackage{amssymb} 	%% The amssymb package provides various useful mathematical symbols

\usepackage{color}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{fullpage}
\usepackage{algorithmic}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}
\algsetup{linenosize=\small}
\usepackage[table,xcdraw]{xcolor}
\usepackage{multirow}
\usepackage[super]{nth}
\usepackage{graphicx,adjustbox}
\usepackage{caption}
\usepackage[labelformat=simple]{subcaption}
\usepackage{comment}
\usepackage{setspace}
\usepackage{textcomp}
\usepackage{xspace}
\usepackage{siunitx}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{soul}
\usepackage{url}
\usepackage{tablefootnote}
\DeclareMathOperator{\E}{\mathbb{E}} % Expectation Symbol
\usepackage[linesnumbered,ruled]{algorithm2e}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage[normalem]{ulem}
\usepackage{footnote}
\usepackage[misc,geometry]{ifsym} 
\makesavenoteenv{tabular}

@@

\renewcommand\thesubfigure{(\alph{subfigure})}
\let\OldTexttrademark\texttrademark
\renewcommand{\texttrademark}{\OldTexttrademark\xspace}%
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

%\usepackage{authblk}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}
\usepackage{epsfig}
%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}
\newcommand{\francesc}[1]{\textcolor{blue}{#1}}
\newcommand{\francesco}[1]{\textcolor{orange}{#1}}
%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}

\journal{Computer Networks}

@@

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

@@

\author[label1, label4]{Francesc Wilhelmi}\ead{francisco.wilhelmi@upf.edu}
\author[label2]{Cristina~Cano}
\author[label3]{Gergely~Neu}
\author[label1]{Boris~Bellalta} 
\author[label3]{Anders~Jonsson}
\author[label1]{Sergio~Barrachina-Mu\~noz}
\address[label1]{Wireless Networking Research Group (WN-UPF), 08002 Barcelona, Spain}
\address[label2]{Wireless Networks Research Group (WINE-UOC), 08860 Castelldefels (Barelona), Spain}
\address[label3]{Artificial Intelligence and Machine Learning Research Group (AIML-UPF), 08002 Barcelona, Spain}

\begin{abstract}
Next-generation wireless deployments are characterized by being dense and uncoordinated, which often leads to inefficient use of resources and poor performance. To solve this, we envision the utilization of completely decentralized mechanisms to \francesc{enable} Spatial Reuse (SR). In particular, we concentrate in Reinforcement Learning (RL), and more specifically, in Multi-Armed Bandits (MABs), to allow networks to modify both their transmission power and channel based on their experienced throughput. In this work, we study the exploration-exploitation trade-off by means of the $\varepsilon$-greedy, EXP3, UCB and Thompson sampling action-selection strategies. Our results show that optimal proportional fairness can be achieved, even if no information about neighboring networks is available to the learners and WNs operate selfishly. However, there is high temporal variability in the throughput experienced by the individual networks, specially for $\varepsilon$-greedy and EXP3. We identify the cause of this variability to be the adversarial setting of our setup in which the set of most played actions provide intermittent good/poor performance depending on the neighboring decisions. We also show that this variability is reduced using UCB and Thompson sampling, which are parameter-free policies that perform exploration according to the reward distribution of each action.
\end{abstract}

\begin{keyword}
High-Density Wireless Networks, Interference, Decentralized learning, Multi Armed Bandits, Spatial Reuse, Resource Allocation.
\end{keyword}

\end{frontmatter}

%\vfill
%Francesc Wilhelmi (e-mail \href{francisco.wilhelmi@upf.edu}{francisco.wilhelmi@upf.edu}), Sergio Barrachina-Mu\~noz and Boris Bellalta are with the Wireless Networking Research Group, Universitat Pompeu Fabra, 08002 Barcelona, Spain. Cristina Cano is with the Wireless Networks Research Group, Universitat Oberta de Catalunya (UOC), 08860 Castelldefels (Barelona), Spain. Anders~Jonsson and Gergely~Neu are with the Artificial Intelligence and Machine Learning Research Group, Universitat Pompeu Fabra (UPF), 08002 Barcelona, Spain.

\newpage
%% \linenumbers

%%%%%%%%%%%%%%%%%%%%%%%
%%%  I. INTRODUCTION 
%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{section:introduction}
Due to the growing popularity of wireless deployments, especially the ones based on the IEEE 802.11 standard (i.e., Wi-Fi), it is very common to find independent overlapping Wireless Networks (WNs) \francesc{sharing the same channel resources. Due to the decentralized nature of such kind of deployments, the lack of organization and/or agreement on sharing policies leads to inefficient resources allocation, both in spectral and spatial domains. An illustrative example about the inefficiency noticed in WNs regarding channel sharing can be found in \cite{akella2007self}, where the authors show that the power level used by wireless devices is typically set, by default, to the maximum, regardless of the distance between communicating nodes, and of the channel occupancy \cite{akella2007self}. In decentralized environments}, providing a model to obtain the optimal solution can be extremely complex. In WNs, the interactions between devices depend on many features (such as position, environment, transmit power) and are hard to derive. On top of this, network dynamics, in terms of user arrivals and departures, makes the resource allocation problem even harder.

@@

From the family of online algorithms, we propose \francesc{the utilization of} Multi-Armed Bandits (MABs) \cite{BCB12}, which is a well-known model in the online learning literature for solving resource allocation problems. In MABs, \francesc{a given agent seeks} to learn a hidden reward distribution while maximizing the gains. This is known as the exploration-exploitation trade-off. %(exploitation focuses on obtaining the maximum instantaneous reward, while exploration aims to maximize the long-term reward)	
While exploitation selects actions with the goal of maximizing long-term reward given the current estimate, exploration selects actions with the goal of improving the estimate. \francesc{Unlike for classical RL, MABs do not consider states\footnote{\francesc{A state refers to a particular situation experienced by a given agent, which is defined by a set of conditions. By having an accurate knowledge on its current situation, an agent can define state-specific strategies that maximize its profits.}} in general, which can be hard to define for the decentralized SR problem. On the one hand, spatial interference cannot be binary treated, thus leading to complex interactions between nodes. On the other hand, the adversarial setting unleashed by decentralized deployments increases the system complexity, since the state not only depends on the actions taken by a given node, but also on the adversaries behavior.}

This work extends our previous results presented in \cite{wilhelmi2017implications}, which were obtained by applying a stateless variation of Q-learning to the SR problem in WNs. Here we generalize the contributions done by implementing several action-selection strategies to allow WNs choose the channel and transmit power to use based on the resulting performance. \francesc{In this paper, then, we provide a tutorial-based application of MABs to the decentralized SR problem, in order} to maximize performance while observing the result from using different SR settings. \francesc{According to that, we aim to illustrate the main benefits and drawbacks of applying different well-known learning techniques in a selfish way, i.e., independent WNs base their strategy on their own experienced performance. On the one hand, we evaluate the impact of varying parameters intrinsic to the proposed algorithms on the resulting throughput and fairness. In addition, we analyze the effects of learning selfishly, in order to shed some light on the future of decentralized approaches.} Notably, we observe that even though players act selfishly, there are algorithms %learn to play collaborative actions
learn to play actions that enhance the overall performance, some times at the cost of high temporal variability. Considering selfish WNs and still obtaining collaborative behaviors is appealing to typical chaotic and dynamic deployments as we get rid of the need for message passing, or inference of neighboring conditions. 
%\francesco{As later shown, this condition holds as long as adversarial WNs are able to compete for channel resources on equal terms. Otherwise, selfish learning would lead to unfairness situations, since WNs in a dominant position would take advantage of their situation, thus harming the performance of WNs in weaker situations.}
%	In particular, we exploit dynamic \francesc{channel selection} and transmit power adjustment \francesc{for dealing} with the performance issues resulting from uncoordinated resource sharing in WNs. While a proper frequency planning allows to reduce the interference between wireless devices, tuning the transmit power adds an extra level of SR that can result in improved throughput and fairness. %We aim to adjust both the channel allocation and transmit power by using local information only. This consideration models typical chaotic and dynamic deployments (e.g., residential buildings) without extra overhead (such as information sharing, additional software and/or hardware). 	
%	We consider selfish WNs that aim to maximize their own throughput and evaluate the impact of this on the resulting throughput and fairness while varying parameters intrinsic to the algorithms.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%
%%%  II. RELATED WORK 
%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
\label{section:related_work} 		
\francesc{In this paper we approach the decentralized SR problem through Dynamic Channel Allocation (DCA) and Transmit Power Control (TPC). While the former aims to allocate the available frequency range among potentially overlapping WNs, the latter attempts to increase the number of parallel transmissions by adjusting the power transmitted by each WN. While DCA operates at the frequency level, TPC deals with more complex spatial interactions. For that, an approach that dynamically tunes the transmit power may severely affect higher communication layers such as network (routing) or transport (congestion). TPC directly affects the transmission range, the data rate and the energy consumption, and it may create unidirectional links that unleash fairness issues. In fact, acknowledgments (ACKs) and RTS/CTS frames assume bidirectionality.}

@@

@@

To the best of our knowledge, there is very little related work on applying multi-armed bandit techniques to the problem of resource allocation in WNs. \francesc{In \cite{coucheney2015multi}, the authors propose modeling a resource allocation problem in LTE networks through MABs. In particular, a set of Base Stations (BS) learn the best configuration of Resource Blocks (RBs) in a decentralized way. For that purpose, a variation of EXP3 (so-called Q-EXP3) is proposed, which allows to reduce the strategy set. Despite a regret bound is provided, it is subject to the fact that an optimal resource allocation exists, i.e., every BS obtains the necessary resources. In addition, a large number of iterations is required to find the optimal solution in a relatively small scenario, thus revealing the difficulties shown by decentralized settings.}

\francesc{More related to the problem proposed here,} the authors in \cite{maghsudi2015joint} show a channel selection and power control approach in infrastructureless networks, which is modeled through bandits. In particular, two different strategies are provided to improve the performance of two Device to Device (D2D) users (each one composed by a transmitter and a receiver), which must learn the best channel and transmit power to be selected. Similarly to our problem, users do not have any knowledge on the channel or the other's configuration, so they rely on the experienced performance in order to find the best configuration. An extension of \cite{maghsudi2015joint} is provided by the same authors in \cite{maghsudi2015channel}, which includes a calibrated predictor (referred in the work as \textit{forecaster}) to infer the behavior of the other devices in order to counter act their actions. In each agent, the information of the forecaster is used to choose the highest-rewarding action with a certain probability, while the rest of actions are randomly selected. Henceforth, assuming that all the networks use a given strategy $\mathcal{X}$, fast convergence is ensured. Results show that channel resources are optimally distributed in a very short time frame through a fully decentralized algorithm that does not require any kind of coordination. Both aforementioned works rely in the existence of a unique Nash Equilibrium, which favors convergence. In contrast, in this article we aim to extend Bandits utilization to denser deployments, and, what is more important, to scenarios with limited available resources in which there is not a unique Nash Equilibrium (NE) that allows fast-convergence. Thus, we aim to capture the effects of applying selfish strategies in a decentralized way (i.e., agent $i$ follows a strategy $\mathcal{X}_i$ that does not consider the strategies of the others) and we also provide insight about the importance of past information for learning in dense WNs, which has not been studied before.

@@

\francesc{For simplicity, we consider that each WN is composed by an AP and a single STA, so that downlink traffic is held. Note that in typical uncoordinated wireless deployments (e.g., residential buildings), STAs are typically close to the AP to which they are associated. Thus, having several STAs associated to the same AP does not significantly impact the inter-WNs interference we study in this work.} According to that, $\Gamma_{i}^*$ is computed as $B \cdot \log_{2}(1 + \text{SNR}_{i})$, where $B$ is the bandwidth, and $\text{SNR}_{i}$ is the Signal-to-Noise ratio experienced by the STA of $\text{WN}_i$. We consider this model because it allows to properly capture the interference generated between WNs, which is sought to be minimized.

@@

For the remainder of this work, \francesc{we study the interactions between several WNs placed in a 3-D scenario that occur when applying MABs in a decentralized manner} (with parameters described later in Section \ref{section:simulation_parameters}). For simplicity, we consider WNs to be composed by an Access Point (AP) transmitting to a single Station (STA) in a downlink manner. \francesc{Additionally, to analyze the adversarial setting when WNs share scarce resources,} we consider that the number of channels is equal to half the number of coexisting WNs, thus making the SR even more challenging.

@@

%------------------------
% Simulation Parameters
%------------------------
\subsection{Simulation Parameters}
\label{section:simulation_parameters}
According to \cite{bellalta2016ax}, which provides an overview of the IEEE 802.11ax-2019 standard, a typical high-density scenario for residential buildings contains $0.0033 \text{APs}/\text{m}^3$ (i.e., 100 APs in a $100 \times 20 \times 15$ m area). Accordingly, for simulation purposes, we define a map scenario with dimensions $10\times5\times10$ m, containing from 2 to 8 APs. In addition, for the first part of the simulations, we consider a setting containing 4 WNs that form a grid topology. In it, STAs are placed at the maximum possible distance from the other networks.

>>>>

\documentclass[10pt,journal,compsoc]{IEEEtran}

\ifCLASSOPTIONcompsoc
% IEEE Computer Society needs nocompress option
% requires cite.sty v4.0 or later (November 2003)
\usepackage[nocompress]{cite}
\else
% normal IEEE
\usepackage{cite}
\fi

\usepackage[table,xcdraw]{xcolor}
\usepackage{color}
\usepackage{amsmath}
\usepackage{mathtools}
%\usepackage{fullpage}
\usepackage[table,xcdraw]{xcolor}
\usepackage{multirow}
\usepackage[super]{nth}
\usepackage{comment}
\usepackage{setspace}
\usepackage{textcomp}
\usepackage{xspace}
\usepackage{siunitx}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{soul}
\usepackage{url}
\usepackage{tablefootnote}
\usepackage{booktabs}
\usepackage[misc]{ifsym}
\usepackage{bbding}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{authblk}

\usepackage{caption}
\usepackage{latexsym}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{cprotect}

%% The graphicx package provides the includegraphics command.
\usepackage[]{graphicx}

\usepackage{algorithm}% http://ctan.org/pkg/algorithm
%\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx

\usepackage{caption}
\usepackage{subcaption}

\usepackage[algo2e]{algorithm2e}

%\usepackage{lmodern}

\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator{\E}{\mathbb{E}} % Expectation Symbol

@@

\newcommand{\francesc}[1]{\textcolor{blue}{#1}}

\makeatletter
\newcommand{\removelatexerror}{\let\@latex@error\@gobble}
\makeatother

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

@@

@@

	\author{\IEEEauthorblockN{
		Francesc~Wilhelmi\Envelope,
		Cristina~Cano,
		Gergely~Neu, 
		Boris~Bellalta,
		Anders~Jonsson and
		Sergio~Barrachina-Mu\~noz}}

%	\thanks{F. Wilhelmi, B. Bellalta and S. Barrachina-Mu\~noz are with the Wireless Networking Research Group (WNRG) at Universitat Pompeu Fabra (UPF), Barcelona, 08018 Spain e-maill: francisco.wilhelmi@upf.edu.}
%	\thanks{C. Cano is with the Wireless Networks Research Group (WINE) at Universitat Oberta de Catalunya (UOC), Castelldefels, 08060 Spain.}
%	\thanks{G. Neu and A. Jonsson are with the Artificial Intelligence and Machine Learning Research Group (AIML) at Universitat Pompeu Fabra (UPF), Barcelona, 08018 Spain.}
	%\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}

	% The paper headers
	\markboth{}%
	{Francesc Wilhelmi \MakeLowercase{\textit{et al.}}: Collaborative Spatial Reuse in Wireless Networks via Selfish Multi-Armed Bandits}

	\IEEEtitleabstractindextext{%
		\begin{abstract}
			Next-generation wireless deployments are characterized by being dense and uncoordinated, which often leads to inefficient use of resources and poor performance. To solve this, we envision the utilization of completely decentralized mechanisms to \francesc{enable} Spatial Reuse (SR). In particular, we concentrate in Reinforcement Learning (RL), and more specifically, in Multi-Armed Bandits (MABs), to allow networks to modify both their transmission power and channel based on their experienced throughput. In this work, we study the exploration-exploitation trade-off by means of the $\varepsilon$-greedy, EXP3, UCB and Thompson sampling action-selection strategies. Our results show that optimal proportional fairness can be achieved, even if no information about neighboring networks is available to the learners and WNs operate selfishly. However, there is high temporal variability in the throughput experienced by the individual networks, specially for $\varepsilon$-greedy and EXP3. We identify the cause of this variability to be the adversarial setting of our setup in which the set of most played actions provide intermittent good/poor performance depending on the neighboring decisions. We also show that this variability is reduced using UCB and Thompson sampling, which are parameter-free policies that perform exploration according to the reward distribution of each action.
			%Next-generation wireless deployments are characterized by being dense and uncoordinated. The lack of self-adjustment mechanisms often leads to inefficient use of resources and poor performance. To solve this, we envision the utilization of completely decentralized mechanisms that enhance Spatial Reuse (SR). In particular, we concentrate in Reinforcement Learning (RL), and more specifically, in Multi-Armed Bandits (MABs), to allow networks to modify both their transmission power and channel based on their experienced throughput. In this work, we study the exploration-exploitation trade-off by means of the $\varepsilon$-greedy, EXP3, UCB and Thompson sampling action-selection strategies. Our results show that optimal proportional fairness can be achieved, even if no information about neighboring networks is available to the learners. With that, we prove collaborative behaviors that favor a set of WNs can be obtained even if they operate in a selfish way, which is useful to avoid centralization costs (e.g., message passing, additional hardware). However, there is high temporal variability in the throughput experienced by the individual networks, specially for $\varepsilon$-greedy and EXP3. We identify the cause of this variability to be the adversarial setting of our setup in which the set of most played actions provide intermittent good/poor performance depending on the neighboring decisions. We also show that this variability is reduced using UCB and Thompson sampling, which are parameter-free policies that perform exploration according to the reward distribution of each action. In particular, Thompson sampling shows promising results regarding the achieved performance while keeping temporal variability low.
		\end{abstract}

		% Note that keywords are not normally used for peerreview papers.
		\begin{IEEEkeywords}
			High-Density Wireless Networks, Interference, Decentralized learning, Multi Armed Bandits, Resource Allocation.
		\end{IEEEkeywords}}

	\maketitle

	\IEEEdisplaynontitleabstractindextext

	\IEEEpeerreviewmaketitle

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	% I. INTRODUCTION											 %
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}

	\IEEEPARstart{D}{ue} to the growing popularity of wireless deployments, especially the ones based on the IEEE 802.11 standard (i.e., Wi-Fi), it is very common to find independent overlapping Wireless Networks (WNs) \francesc{sharing the same channel resources. Due to the decentralized nature of such kind of deployments, the lack of organization and/or agreement on sharing policies leads to inefficient resources allocation, both in spectral and spatial domains. An illustrative example about the inefficiency noticed in WNs regarding channel sharing can be found in \cite{akella2007self}, where the authors show that the power level used by wireless devices is typically set, by default, to the maximum, regardless of the distance between communicating nodes, and of the channel occupancy \cite{akella2007self}. In decentralized environments}, providing a model to obtain the optimal solution can be extremely complex. In WNs, the interactions between devices depend on many features (such as position, environment, transmit power) and are hard to derive. On top of this, network dynamics, in terms of user arrivals and departures, makes the resource allocation problem even harder.

@@

	From the family of online algorithms, we propose \francesc{the utilization of} Multi-Armed Bandits (MABs) \cite{BCB12}, which is a well-known model in the online learning literature for solving resource allocation problems. In MABs, \francesc{a given agent seeks} to learn a hidden reward distribution while maximizing the gains. This is known as the exploration-exploitation trade-off. %(exploitation focuses on obtaining the maximum instantaneous reward, while exploration aims to maximize the long-term reward)	
	While exploitation selects actions with the goal of maximizing long-term reward given the current estimate, exploration selects actions with the goal of improving the estimate. \francesc{Unlike for classical RL, MABs do not consider states\footnote{\francesc{A state refers to a particular situation experienced by a given agent, which is defined by a set of conditions. By having an accurate knowledge on its current situation, an agent can define state-specific strategies that maximize its profits.}} in general, which can be hard define for the decentralized SR problem. On the one hand, spatial interference cannot be binary treated, thus leading to complex interactions between nodes. On the other hand, the adversarial setting unleashed by decentralized deployments increases the , since the state not only depends on the actions taken by a given node, but also on the adversaries behavior.}

	This work extends our previous results presented in \cite{wilhelmi2017implications}, which were obtained by applying a stateless variation of Q-learning to the SR problem in WNs. Here we generalize the contributions done by implementing several action-selection strategies to allow WNs choose the channel and transmit power to use based on the resulting performance. \francesc{In this paper, then, we provide a tutorial-based application of MABs to the decentralized SR problem, in order} to maximize performance while observing the result from using different SR settings. \francesc{According to that, we aim to illustrate the main benefits and drawbacks of applying different well-known learning techniques in a selfish way, i.e., independent WNs base their strategy on their own experienced performance. On the one hand, we evaluate the impact of varying parameters intrinsic to the proposed algorithms on the resulting throughput and fairness. In addition, we analyze the effects of learning selfishly, in order to shed some light on the future of decentralized approaches.} 	Notably, we observe that even though players act selfishly, there are algorithms %learn to play collaborative actions
	learn to play actions that enhance the overall performance, some times at the cost of high temporal variability. Considering selfish WNs and still obtaining collaborative behaviors is appealing to typical chaotic and dynamic deployments as we get rid of the need for message passing, or inference of neighboring conditions.

%	In particular, we exploit dynamic \francesc{channel selection} and transmit power adjustment \francesc{for dealing} with the performance issues resulting from uncoordinated resource sharing in WNs. While a proper frequency planning allows to reduce the interference between wireless devices, tuning the transmit power adds an extra level of SR that can result in improved throughput and fairness. %We aim to adjust both the channel allocation and transmit power by using local information only. This consideration models typical chaotic and dynamic deployments (e.g., residential buildings) without extra overhead (such as information sharing, additional software and/or hardware). 	
%	We consider selfish WNs that aim to maximize their own throughput and evaluate the impact of this on the resulting throughput and fairness while varying parameters intrinsic to the algorithms.

@@

	%------------------------
	% Article Structure
	%------------------------

@@

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	% II. RELATED WORK											 %
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\section{Related Work}
	\label{section:related_work}	
	\francesc{In this paper we approach the decentralized SR problem through Dynamic Channel Allocation (DCA) and Transmit Power Control (TPC). While the former aims to allocate the available frequency range among potentially overlapping WNs, the latter attempts to increase the number of parallel transmissions by adjusting the power transmitted by each WN. While DCA operate at the frequency level, TPC deals with more complex spatial interactions. For that, an approach that dynamically tunes the transmit power may severely affect higher communication layers such as network (routing) or transport (congestion). TPC directly affects the transmission range, the data rate and the energy consumption, and it may create unidirectional links that unleash fairness issues. In fact, acknowledgments (ACKs) and RTS/CTS frames assume bidirectionality.}

@@

%	Dynamic Channel Allocation (DCA) aims to allocate the available frequency range among overlapping wireless devices, so that interference is reduced and the aggregate throughput is maximized. We find many centralized solutions based on graph coloring techniques \cite{riihijarvi2005frequency, mishra2005weighted}, which result to be impractical in uncoordinated scenarios, since there is no communication between the implied nodes.

	%The authors in \cite{akl2007dynamic} propose a very simple approach in which each AP maintains an interference map of their neighbors , such that channel assignment is done through interference minimization. Unfortunately, the interactions between APs due to decentralization are not studied. Separately, \cite{chen2007improved} proposes two decentralized approaches and one centralized that rely in the interference measured at both APs and STAs to calculate the best frequency channels for dynamic channel allocation. To do so, a WN, in addition to the interference sensed by its associated devices, considers other metrics such as the amount of traffic, so that some kind of coordination is required at the neighbor level (periodic reporting). With respect to the centralized approach, which is shown to significantly improve the decentralized solutions, the main drawback is the requirement of a wired infrastructure for the management communication. Such deployment is often infeasible, especially in typical residential scenarios.

%	Regarding Transmit Power Control (TPC), it is possible to increase the number of parallel transmissions or reaching farther distances by adjusting the power transmitted by each WN. However, modifying the transmit power may severely affect higher communication layers such as network (routing) or transport (congestion). TPC directly affects the transmission range, the data rate and the energy consumption, and it may create unidirectional links that unleash fairness issues. In fact, acknowledgments (ACKs) and RTS/CTS frames assume bidirectionality. One of the first works in studying power management is \cite{elbatt2000power}, which shows that large improvements can be achieved if selecting the appropriate transmit power level in ad-hoc WNs. A centralized approach for TPC is presented in \cite{tang2014joint}, which performs power control and Rate Adaptation (RA) in subgroups of Wireless Local Area Networks (WLANs). The creation of clusters allows defining independent power levels between devices in the same group, which are useful to avoid asymmetric links. However, to represent all the possible combinations, graphs can become very large, specially in high-density deployments. Furthermore, we find a decentralized approach that relies on real-time channel measurements \cite{gandarillas2014dynamic}. The proposed mechanism (so called Dynamic Transmission Power Control) is based on a set of triggered thresholds that increase/decrease the transmit power according to the situation. The main problem is that thresholds are set empirically (based on simulations), so the mechanism may not always improve performance.

@@

	To the best of our knowledge, there is very little related work on applying multi-armed bandit techniques to the problem of resource allocation in WNs. \francesc{In \cite{coucheney2015multi}, the authors propose modeling a resource allocation problem in LTE networks through MABs. In particular, a set of Base Stations (BS) learn the best configuration of Resource Blocks (RBs) in a decentralized way. For that purpose, a variation of EXP3 (so-called Q-EXP3) is proposed, which allows to reduce the strategy set. Despite a regret bound is provided, it is subject to the fact that a perfect resource allocation exists. In addition, a large number of iterations is required to find the optimal solution in a relatively small scenario, thus revealing the difficulties shown by decentralized settings.}

	\francesc{With a higher degree of relation to the problem proposed here,} the authors in \cite{maghsudi2015joint} show a channel selection and power control approach in infrastructureless networks, which is modeled through bandits. In particular, two different strategies are provided to improve the performance of two Device to Device (D2D) users (each one composed by a transmitter and a receiver), which must learn the best channel and transmit power to be selected. Similarly to our problem, users do not have any knowledge on the channel or the other's configuration, so they rely on the experienced performance in order to find the best configuration. An extension of \cite{maghsudi2015joint} is provided by the same authors in \cite{maghsudi2015channel}, which includes a calibrated predictor (referred in the work as \textit{forecaster}) to infer the behavior of the other devices in order to counter act their actions. In each agent, the information of the forecaster is used to choose the highest-rewarding action with a certain probability, while the rest of actions are randomly selected. Henceforth, assuming that all the networks use a given strategy $\mathcal{X}$, fast convergence is ensured. Results show that channel resources are optimally distributed in a very short time frame through a fully decentralized algorithm that does not require any kind of coordination. Both aforementioned works rely in the existence of a unique Nash Equilibrium, which favors convergence. In contrast, in this article we aim to extend Bandits utilization to denser deployments, and, what is more important, to scenarios with limited available resources in which there is not a unique Nash Equilibrium (NE) that allows fast-convergence. Thus, we aim to capture the effects of applying selfish strategies in a decentralized way (i.e., agent $i$ follows a strategy $\mathcal{X}_i$ that does not consider the strategies of the others) and we also provide insight about the importance of past information for learning in dense WNs, which has not been studied before.

@@

	\francesc{For simplicity, we consider that each WN is composed by an AP and a single STA. Note that in typical uncoordinated wireless deployments (e.g., residential buildings), STAs are typically close to the AP to which they are associated. Thus, having several STAs associated to the same AP does not significant impact to the inter-WNs interference that is aimed to be studied in this work.} According to that, $\Gamma_{i}^*$ is computed as $B \cdot \log_{2}(1 + \text{SNR}_{i})$, where $B$ is the bandwidth, and $\text{SNR}_{i}$ is the Signal-to-Noise ratio experienced by the STA of $\text{WN}_i$. We consider this model because it allows to properly capture the interference generated between WNs, which is sought to be minimized.

@@

	For the remainder of this work, we consider several WNs placed in a 3-D scenario in order to study the effects of applying MABs in a decentralized manner (with parameters described later in Section \ref{section:simulation_parameters}). For simplicity, we consider WNs to be composed by an Access Point (AP) transmitting to a single Station (STA) in a downlink manner. We consider as well that the number of channels is equal to half the number of coexisting WNs, so that SR is challenging.

@@

	%------------------------
	% Simulation Parameters
	%------------------------
	\subsection{Simulation Parameters}
	\label{section:simulation_parameters}
	According to \cite{bellalta2016ax}, which provides an overview of the IEEE 802.11ax-2019 standard, a typical high-density scenario for residential buildings contains $0.0033 \text{APs}/\text{m}^3$ (i.e., 100 APs in a $100 \times 20 \times 15$ m area). Accordingly, for simulation purposes, we define a map scenario with dimensions $10\times5\times10$ m, containing from 2 to 8 APs. In addition, for the first part of the simulations, we consider a setting containing 4 WNs that form a grid topology. In it, STAs are placed at the maximum possible distance from the other networks. The 4-grid scenario is illustrated in Figure \ref{fig:scenario}.
	\begin{figure}[t!]
		\centering											
		\epsfig{file=images/4_WLANs_scenario.eps, width=6.5cm}
		\caption{Grid scenario containing 4 WNs, each one composed by an AP and a STA.}
		\label{fig:scenario}
	\end{figure}
	Table \ref{tbl:simulation_parameters} details the parameters used.	
	\begin{table}[h!]
		\centering
		\resizebox{\columnwidth}{!}{
			\begin{tabular}{|l|l|}
				\hline
				\textbf{Parameter}             & \textbf{Value}                      \\ \hline
				Map size (m)                    & $10\times5\times10$                  \\ \hline
				Number of coexistent WNs      & \{2, 4, 6, 8\} \\ \hline
				APs/STAs per WN                    & 1 / 1                                     \\ \hline
				Distance AP-STA (m)     & $\sqrt{2}$                         \\ \hline
				Number of Channels              & 2                               \\ \hline
				Channel Bandwidth (MHz)         & 20                                    \\ \hline
				Initial channel selection model & Uniformly distributed 
				\\ \hline
				TPC Values (dBm)                & \{5, 10, 15, 20\}                      \\ \hline				
				$\text{PL}_0$ (dB)                   & 5
				\\ \hline
				$\text{G}_s$ (dB)                   & Normally distributed with mean 9.5 
				\\ \hline
				$\text{G}_o$ (dB)                  & Uniformly distributed with mean 30
				\\ \hline
				$d_{\rm obs}$  (meters between two obstacles)                 & 5                        
				\\ \hline
				Noise level (dBm)               & -100                                  \\ \hline
				Traffic model                   & Full buffer (downlink)             \\ \hline          
				Number of learning iterations                  & 10,000             \\ \hline   
		\end{tabular}}
		\caption{Simulation parameters}
		\label{tbl:simulation_parameters}
	\end{table}
